---
title: "Overview Summary"
author: "TEnright"
date: "11/08/2019"
output:
  html_document:
    df_print: paged
    toc: yes
params:
  optimal_res: "optimal_res.txt"
  destDir: "PATH/TO/OUTPUT"
  seed: "1234"
  GOdown: "GOdown.txt"
  GOup: "GOup.txt"
  type: "RNA-or-integrated"
  markers: "markers_file.txt"
  contrast: "condition"
  naive_seurat: "seurat.rds"
  output_file: "path/to/output"
---

1. Read nomalized,integrated, and labelled seurat object
2. Cell/Cluster Information
3. Differential Expression

```{r setup_libs, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(data.table)
library(knitr)
library(ggplot2)
library(grid)
library(gridExtra)
library(DT)
library(plyr)
library(future)
library(tidyr)
library(plotly)
options(future.globals.maxSize = 96000 * 1024^2)
set.seed(as.numeric(params$seed))
```

# 1. Read nomalized and integrated seurat object
```{r readObject, warning = F, echo = T}
naive_seurat <- params$naive_seurat
output_file <- params$output_file
out_dir <- params$destDir
type <- params$type
marker_file <- params$markers
GO_down <- params$GOdown
GO_up <- params$GOup
optimal_res_file <- params$optimal_res
contrast <- params$contrast

if(!(file.exists(out_dir))) {
  dir.create(out_dir,showWarnings = FALSE, recursive = TRUE)
}



optimal_res <- as.double(read.table(optimal_res_file)[1,1]) #according to the python script from Joey
assay <- "integrated"

#load seurat object
input_dir <- paste(strsplit(optimal_res_file,"/")[[1]][-length(strsplit(optimal_res_file,"/")[[1]])],collapse="/")
input_rds <- grep(paste(sprintf("^%s",optimal_res),"SeuratObj.rds",sep="_"),list.files(input_dir),value = T)[1]
seuratObj = readRDS(paste(input_dir,input_rds,sep="/"))

seuratObj@misc[["optimal_res"]] <- optimal_res

```


```{r helper_functions, echo = F, eval = T}


recursive_paste <- function(object,list_names,sep = "_"){
  length_n <- length(list_names)
  if ((length_n) == 1){
    return(object@meta.data[[list_names[1]]])
  }
	if ((length_n) == 2){
		return(paste(object@meta.data[[list_names[length_n - 1]]],object@meta.data[[list_names[length_n]]],sep = sep))
	}
	else{
		paste_1 <- list_names[length_n - 1]
		paste_2 <- list_names[length_n]
		pasted <- paste(paste_1,paste_2,sep = sep)
		object <- AddMetaData(object = object, paste(object@meta.data[[paste_1]],object@meta.data[[paste_2]],sep = sep), col.name = pasted)
		reduced_list <- list_names[1:(length_n-2)]
		new_list <- c(reduced_list,pasted)
		return(recursive_paste(object,new_list,sep = sep))
	}
}

get_overlap_combi <- function(df,combi,sep){
  new_list <- NULL
  for (i in 1:length(combi)){
    current_piece <- sapply(strsplit(as.character(df[,1]),sprintf("\\%s",sep)), "[[",combi[i])
    if (i == 1){
      new_list <- current_piece
    }
    else{
    new_list <- paste(new_list,current_piece, sep = sep)
    }
  }
  return(new_list)
}

get_new_column <- function(df,names_list,combi,sep){
  column_new <- NULL
  df <- get_overlap_combi(df,combi,sep)
  # loop through and get 
  for (i in names_list){
    locations <- grep(i,df)
    column_new[locations] <- i
  }
  return(column_new)
}

make_meta_counts_table <- function(object,meta_list,sep){
  #requires library tidyr
  # and dpylr
  #object = seurat Object
  #meta_list = column names from meta.data
  #sep = the separator to use. This is fragile depending on how the data in the columns are parsed. for example if sep = "_" and there are underscores in the meta data then this function will behave erradictly or break completely
	#get starting data frame
	df <- as.data.frame(table(recursive_paste(object,meta_list,sep = sep)))
	#go through different combinations starting with groups of n, then n-1, ... then 1, where n is equal to the length of the meta_list
	for (i in 1:(length(meta_list)-1)){
	if(i > length(meta_list)-1){break}
  matrix_comb <- combn(meta_list,i) #cominations with "ith" number of meta columns. ie:  if i = 2, then matrix_comb is all combinations of the meta_list ((length of meta list) CHOOSE 2 combinations)
    for (k in 1:ncol(matrix_comb)){
        current_combination <- matrix_comb[,k]
        tab <- table(recursive_paste(object,current_combination,sep= sep))
        n_combi <- which(meta_list %in% current_combination)
        new_col_name <- paste(matrix_comb[,k],collapse = sep)
        df[[new_col_name]] <- get_new_column(df,unique(names(tab)),n_combi,sep)
        df[[paste("n_cells",new_col_name,sep = sep)]] <- mapvalues(df[[new_col_name]],from = names(tab),to = tab)
    }
  }
	return(df)
}

```


# 2. Cell/Cluster Identification




## Bulk

avg_logFC: log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group

*pct.1*: The percentage of cells where the gene is detected in the *contrast*

*pct.2*: The percentage of cells where the gene is detected in the *baseline*

p_val_adj: Adjusted p-value, based on bonferroni correction using all genes in the dataset


```{r echo = F, message=F, warning=F}


n_genes <- 10


bulk_DE <- read.delim(marker_file,header=TRUE,row.names=1,sep="\t")
#seuratObj$condition2 <- mapvalues(seuratObj$condition, from = "flow", to = "fresh")

datatable(bulk_DE)
```


```{r echo = F, eval = T, warning = F, message = F}

df <- make_meta_counts_table(seuratObj,meta_list = unique(seuratObj@misc[["meta_tab"]]),sep = "_")
colnames(df) <- c("SampleID","n_Cells",colnames(df)[3:length(colnames(df))])
kable(df)
```

```{r echo  = F, eval = T, cache=T}


bulk_DE$delta <- log2(bulk_DE$pct.1/bulk_DE$pct.2)
bulk_DE$delta_pc <- bulk_DE$pct.1-bulk_DE$pct.2

bulk_DE_filtered <- subset(bulk_DE, subset = p_val < 0.05)

bulk_DE_filtered$label <- paste(rownames(bulk_DE_filtered)," : ", bulk_DE_filtered$avg_logFC, sep="")
plot_ly(bulk_DE_filtered,
          x = ~avg_logFC, y = ~delta_pc,
          color = ~p_val_adj,
           # you can just run this against the column for the gene as well using ~ACTB, the algorith will automatically scale in that case based on maximal and minimal values
          opacity = .5,
          type = "scatter", 
          mode = "markers",
          marker = list(size = 5, width=2), 
          text=~label,
          hoverinfo="text"
)

```

## UMAP

## Naive

```{r naive, echo = F, eval = T}


naive_seurat <- readRDS(naive_seurat)
dimP <- DimPlot(object = naive_seurat, reduction = "umap", label = F, group.by = "orig.ident") + theme(text = element_text(size = 8))
dimP <- LabelClusters(plot = dimP,id = "orig.ident",repel = T, size = 2.4)
dimP

```


## Integration with anchoring cells via seurat Integration

```{r dimPlot_umaps1, echo = F, eval = T}

dimP <- DimPlot(object = seuratObj, reduction = "umap", label = F, group.by = "orig.ident") + theme(text = element_text(size = 8))
dimP <- LabelClusters(plot = dimP,id = "orig.ident",repel = T, size = 2.4)

dimP

```

```{r dimPlot_umaps, echo = F, eval = T}
column_cluster <- sprintf("integrated_snn_res.%s",optimal_res)
dimP <- DimPlot(object = seuratObj, reduction = "umap", label = F, group.by = column_cluster) + theme(text = element_text(size = 8))
dimP <- LabelClusters(plot = dimP,id = column_cluster,repel = T, size = 2.4)
ggsave(filename = sprintf(fmt = "%s_%s_optimal_cluster.tiff",seuratObj@project.name,optimal_res), plot = dimP, device = "tiff",path = out_dir)
dimP

```

```{r dimPlot_umaps_split, echo = F, eval = T}



for (meta in seuratObj@misc[["meta_tab"]]){
	dimP <- DimPlot(object = seuratObj, reduction = "umap", label = F, group.by = column_cluster, split.by = meta) + theme(text = element_text(size = 8))
	dimP <- LabelClusters(plot = dimP,id = column_cluster,repel = T, size = 2.4)
	ggsave(filename = sprintf(fmt = "%s_%s_split_by_umap.tiff",meta,optimal_res), plot = dimP, device = "tiff",path = out_dir)
	print(dimP)
}

```



```{r GO_plots, echo = F, eval = T }
library(ggrepel)
go_down <- read.delim(GO_down)
go_up <- read.delim(GO_up)

go_down <- subset(go_down, subset= p.adj < 0.05)

go_down$enrich[grep("Inf",go_down$enrich)] <- NA
scaled_go <- go_down$enrich/sd(go_down$enrich,na.rm = T)
max_scaled <- max(scaled_go[which(!is.na(scaled_go))])
scaled_go[which(is.na(scaled_go))] <- max_scaled+0.5
go_down$enrich <- scaled_go

p <- ggplot(data = go_down, aes(x=enrich, y = p.adj, size = Annotated, color = Term)) + geom_point(position = "jitter",alpha=0.5) + ggtitle("Down") + geom_label_repel(aes(label=Term),point.padding = 0.5,size = 3.5, show.legend = F)
	
ggplotly(p)



go_up <- subset(go_up, subset= p.adj < 1)

go_up$enrich[grep("Inf",go_up$enrich)] <- NA
scaled_go <- go_up$enrich/sd(go_up$enrich,na.rm = T)
max_scaled <- max(scaled_go[which(!is.na(scaled_go))])
scaled_go[which(is.na(scaled_go))] <- max_scaled+0.5
go_up$enrich <- scaled_go

p <- ggplot(data = go_up, aes(x=enrich, y = p.adj, size = Annotated, color = Term)) + geom_point(position = "jitter",alpha=0.5) + ggtitle("Up") + geom_label_repel(aes(label=Term),point.padding = 0.5,size = 3.5, show.legend = F)
	
ggplotly(p)


```



```{r echo = F, eval = T, warning = F}

#uses cononical correlation analysis and mutual neighest neighbor to determine putative "anchors" 
#CCA first finds the covariance between the datasets and then finds mutual neighest neighbors as putative "anchors"
#Low probability anchors get filitred out before 

meta_tab <- unique(seuratObj@misc[["meta_tab"]])
gpLot <- DimPlot(seuratObj, reduction = "umap", group.by = meta_tab[1])
gbuild <- ggplot2::ggplot_build(gpLot)
colors <- unique(gbuild[[1]][[1]]$colour)




plotting.data <- FetchData(object = seuratObj, vars = c("UMAP_1", "UMAP_2", "UMAP_3", meta_tab[1]), slot = 'data')

plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[[meta_tab[1]]], sep="")


plot_ly(data = plotting.data, 
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = plotting.data[[meta_tab[1]]], # you can just run this against the column for the gene as well using ~ACTB, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = rev(colors), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 5, width=2), 
        text=~label,
        hoverinfo="text"
)


gpLot <- DimPlot(seuratObj, reduction = "umap", group.by = meta_tab[2])
gbuild <- ggplot2::ggplot_build(gpLot)
colors <- unique(gbuild[[1]][[1]]$colour)

plotting.data <- FetchData(object = seuratObj, vars = c("UMAP_1", "UMAP_2", "UMAP_3", meta_tab[2]), slot = 'data')

plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[[meta_tab[2]]], sep="")


plot_ly(data = plotting.data, 
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = plotting.data[[meta_tab[2]]], # you can just run this against the column for the gene as well using ~ACTB, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = rev(colors), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 5, width=2), 
        text=~label,
        hoverinfo="text"
)


gpLot <- DimPlot(seuratObj, reduction = "umap", group.by = meta_tab[3])
gbuild <- ggplot2::ggplot_build(gpLot)
colors <- unique(gbuild[[1]][[1]]$colour)

plotting.data <- FetchData(object = seuratObj, vars = c("UMAP_1", "UMAP_2", "UMAP_3", meta_tab[3]), slot = 'data')

plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[[meta_tab[3]]], sep="")


plot_ly(data = plotting.data, 
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = plotting.data[[meta_tab[3]]], # you can just run this against the column for the gene as well using ~ACTB, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = rev(colors), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 5, width=2), 
        text=~label,
        hoverinfo="text"
)




```


```{r cellBrowser, echo = F, eval = F}
seuratObj$clusters <- seuratObj@meta.data[[sprintf(integrated_snn_res.%s,optimal_res)]]
ExportToCellbrowser(seuratObj,dir='./cellBrowser',dataset.name=seuratObj@project.name,reductions=c('umap'),markers.file=file_name,cluster.field='clusters',Phase='Phase',condition='condition',clusters='clusters')

```

```{r save_RDS, echo = F, eval = T}

saveRDS(seuratObj,file = output_file)


```



#### Session Information
```{r, echo=FALSE}
sessionInfo()
```

